---
title: "Interactive Maps for Study Data"
author: "Norman Hassell"
date: "February 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Maps of Chemical, Biological, and Physical Data

First we'll load data from previous analyses to avoid re-computation.
```{r}
load("C:/Users/Norman/Dropbox/RiverPaper_work/2017_01_08_correction_work/R_projects_for_paper/figure_new_spearman_histograms/.RData")
```

Then load all necessary packages.
```{r}
library(leaflet)
library(colorRamps)
library(maps)
library(rgdal)
library(raster)
library(RColorBrewer)
```

Now add columns of relevant metadata to habitat data-frame to allow mapping
```{r}
#load complete metadata csv
SP13toSP14_meta_complete <- read.csv("/Users/Norman/Dropbox/RiverPaper_work/2017_01_08_correction_work/SP13toSP14_sampleData_02.19.17_complete_nb.csv", stringsAsFactors = F)

#order data
target <- SP13toSP14_20170122_meta$Sample
SP13toSP14_meta_complete_ord <- SP13toSP14_meta_complete[match(target, SP13toSP14_meta_complete$Sample),]

#remove uncessary columns
habitat_otu_tab_ord_ed <- habitat_otu_tab_ord[,-c(11,12)]

#create necessary data columns
Longitude <- SP13toSP14_20170122_meta$Longitude
Latitude <- SP13toSP14_20170122_meta$Latitude
TDN_uM <- SP13toSP14_20170122_meta$TDN_uM
TDP_uM <- SP13toSP14_20170122_meta$TDP_uM
Conductivity <- SP13toSP14_meta_complete_ord$Conductivity_us.cm
Turbidity <- SP13toSP14_meta_complete_ord$Turbidity._NTU
Nitrates <- SP13toSP14_meta_complete_ord$Nitrates._mg.L
pH <- SP13toSP14_meta_complete_ord$pH
E_coli <- SP13toSP14_meta_complete_ord$E_.Coli._CFU.100ml_

Stream_Length_km <- SP13toSP14_20170122_meta$Stream_Length_km


habitat_otu_map <- cbind.data.frame(Longitude, Latitude, TDN_uM, TDP_uM, Conductivity, Turbidity, Nitrates, pH, E_coli, Stream_Length_km, habitat_otu_tab_ord_ed, stringsAsFactors = F)
```

Then add shape-file of total watershed area of the study.
```{r}

watershed <- shapefile("C:/Users/Norman/Dropbox/new_gis_stream/dissovled_area.shp")
```


Total dissolved phosphorus:
```{r, echo=F, warning=F}
habitat_otu_map$season <- factor(habitat_otu_map$season, 
                                          levels = habitat_otu_map$season, ordered = T)

pal_TDP <- colorNumeric(
    palette = matlab.like(100),
    domain = habitat_otu_map$TDP_uM
)

content_TDP <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "TDP (µM):", habitat_otu_map$TDP_uM)



#TDP Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
addProviderTiles("OpenTopoMap") %>%
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", fillColor = ~pal_TDP(TDP_uM), weight = 1, fillOpacity = 0.8, opacity = 0.8, popup = content_TDP
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_TDP, values = ~TDP_uM,
              title = paste(sep = "<br/>", "Total", "Dissolved","Phosphorus"), labFormat = labelFormat(suffix = " µM"), opacity = 1)

```

Total dissolved nitrogen:
```{r, echo=F, warning=F}
pal_TDN <- colorNumeric(
    palette = matlab.like(100),
    domain = habitat_otu_map$TDN_uM
)

content_TDN <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "TDN (µM):", habitat_otu_map$TDN_uM)


#TDN Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", opacity = 0.8, fillColor = ~pal_TDN(TDN_uM), weight = 1, fillOpacity = 0.8, popup = content_TDN
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_TDN, values = ~TDN_uM,
              title = paste(sep = "<br/>", "Total", "Dissolved","Nitrogen"), labFormat = labelFormat(suffix = " µM"), opacity = 1)

```

Total nitrates.
```{r, echo=F, warning=F}
pal_Nitrates <- colorNumeric(
    palette = matlab.like(100),
    domain = habitat_otu_map$Nitrates
)

content_Nitrates <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "Nitrates (mg/L):", habitat_otu_map$Nitrates)


#Nitrates Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", fillColor = ~pal_Nitrates(Nitrates), fillOpacity = 0.8, weight = 1, opacity = 0.8, popup = content_Nitrates
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_Nitrates, values = ~Nitrates,
              title = paste(sep = "<br/>", "Total", "Nitrates"), labFormat = labelFormat(suffix = " mg/L"), opacity = 1)

```

Conductivity.
```{r, echo=F, warning=F}
pal_Conductivity <- colorNumeric(
    palette = matlab.like(100),
    domain = habitat_otu_map$Conductivity
)

content_Conductivity <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "Conductivity (µs/cm):", habitat_otu_map$Conductivity)



#Conductivity Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
addProviderTiles("OpenTopoMap") %>%
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", weight = 1, fillColor = ~pal_Conductivity(Conductivity), fillOpacity = 0.8, opacity = 0.8, popup = content_Conductivity
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_Conductivity, values = ~Conductivity,
              title = "Conductivity", labFormat = labelFormat(suffix = " µs/cm"), opacity = 1)

```

Sample pH:
```{r, echo=F, warning=F}
pal_pH <- colorNumeric(
    palette = rev(matlab.like(100)),
    domain = habitat_otu_map$pH
)

content_pH <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "pH:", habitat_otu_map$pH)


#pH Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", weight = 1, fillColor = ~pal_pH(pH), fillOpacity = 0.8, opacity = 0.8, popup = content_pH
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_pH, values = ~pH,
              title = "Sample pH", labFormat = labelFormat(prefix = "pH "), opacity = 1)

```

Sample *E. coli* counts.
```{r, echo=F, warning=F}
pal_E_coli <- colorNumeric(
    palette = matlab.like(100),
    domain = habitat_otu_map$E_coli
)

content_E_coli <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "E. coli (CFU/ml):", habitat_otu_map$E_coli)


#E_coli Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", weight = 1, opacity = 0.8, fillOpacity = 0.8, fillColor = ~pal_E_coli(E_coli), popup = content_E_coli
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_E_coli, values = ~E_coli,
              title = paste(sep = "<br/>", "E. coli", "Counts"), labFormat = labelFormat(suffix = " CFU/ml"), opacity = 1)

```

Cumulative stream site lengths:
```{r, echo=F, warning=F}
breaks_flow <- c(1,2,3,4,5,6,7,8,9,10,25,100,250,500,2000)
breaks <- c(0,1,breaks_flow*2)

colors <- rev(rainbow(length(breaks), end=4/6))

pal_Length <- colorBin(
    palette = colors, domain = habitat_otu_map$Stream_Length_km, bins = breaks
)

content_Length <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km)


#TDN Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, fillColor = ~pal_Length(Stream_Length_km), color = "black", weight = 1, opacity = 0.8, fillOpacity = 0.8, popup = content_Length
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomleft", pal = pal_Length, values = ~Stream_Length_km,
              title = paste(sep = "<br/>", "Cumulative", "Stream","Length"), labFormat = labelFormat(suffix = " km"), opacity = 1)

```

##Maps of Sample Community Compososition Data

Community Diversity (Shannon):
```{r, echo=F, warning=F}
pal_shannon <- colorNumeric(
    palette =  rev(brewer.pal(100,"Spectral")), domain = habitat_otu_map$shannon
)

content_shannon <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "Shannon Div.:", habitat_otu_map$shannon)


#TDN Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", opacity = 0.8, fillOpacity = 0.8, weight = 1, fillColor = ~pal_shannon(shannon), popup = content_shannon
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_shannon, values = ~shannon,
              title = paste(sep = "<br/>", "Community", "Diversity","(Shannon)"), opacity = 1)

```

Fraction of OTUs originating from freshwater:
```{r, echo=F, warning=F}
pal_fw_otu_percent <- colorNumeric(
    palette =  rev(brewer.pal(100,"Spectral")), domain = habitat_otu_map$fw_otu_percent
)

content_fw_otu_percent <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "Frac. FW OTUs:", habitat_otu_map$fw_otu_percent)


#TDN Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", opacity = 0.8, fillOpacity = 0.8, weight = 1, fillColor = ~pal_fw_otu_percent(fw_otu_percent), popup = content_fw_otu_percent
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_fw_otu_percent, values = ~fw_otu_percent,
              title = paste(sep = "<br/>", "Freshwater", "OTUs"), labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x), opacity = 1)

```

Fraction of OTUs originating from soil:
```{r, echo=F, warning=F}
pal_soil_otu_percent <- colorNumeric(
    palette =  rev(brewer.pal(100,"Spectral")), domain = habitat_otu_map$soil_otu_percent
)

content_soil_otu_percent <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "Frac. Soil OTUs:", habitat_otu_map$soil_otu_percent)


#TDN Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", opacity = 0.8, fillOpacity = 0.8, weight = 1, fillColor = ~pal_soil_otu_percent(soil_otu_percent), popup = content_soil_otu_percent
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_soil_otu_percent, values = ~soil_otu_percent,
              title = paste(sep = "<br/>", "Soil", "OTUs"), labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x), opacity = 1)

```

Fraction of core OTUs:
```{r, echo=F, warning=F}
pal_core_otus <- colorNumeric(
    palette =  rev(brewer.pal(100,"Spectral")), domain = habitat_otu_map$core_otus
)

content_core_otus <- paste(sep = "<br/>", habitat_otu_map$samples, "Cum. Len. (km):",
                     habitat_otu_map$Stream_Length_km, "Frac. Core OTUs:", habitat_otu_map$core_otus)


#TDN Map
leaflet(data = habitat_otu_map) %>%
  addPolygons(data = watershed,
    stroke = T, fillColor = "transparent", color = "black", weight = 2
  ) %>%
    addProviderTiles("OpenTopoMap") %>% 
    addCircleMarkers(
        lat = ~ Latitude, lng = ~ Longitude, radius = 4, group = ~season, color = "black", opacity = 0.8, fillOpacity = 0.8, weight = 1, fillColor = ~pal_core_otus(core_otus), popup = content_core_otus
        ) %>% 
    addLayersControl(overlayGroups = ~season) %>% 
    addLegend("bottomright", pal = pal_core_otus, values = ~core_otus,
              title = paste(sep = "<br/>", "Core", "OTUs"), labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x), opacity = 1)

```